---
title: "Week 5 Lab: Time series data and mapping"
author: "Victoria Wallace"
date: "2/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
```

### **Step 1:** Attach packages.

```{r}
# General:
library(tidyverse)
library(janitor)
library(lubridate)
library(here)
library(paletteer)

# ts: 
library(tsibble)
library(fable)
library(fabletools)
library(feasts)
library(forecast)

# Spatial: 
library(sf)
library(tmap)
library(mapview)
```

### **Step 2:** Read in data on monthly U.S. energy consumption for renewables

```{r}
us_renew <- read_csv(here::here("data", "renewables_cons_prod.csv")) %>% 
  clean_names()
```

Some exploration:

Explore the data frame:

```{r}
# In the console, I've run:
# unique(us_renew$description)
```

I will only be exploring consumption. But first, I need to do some cleaning and wrangling:


```{r}
renew_clean <- us_renew %>%
  mutate(description = str_to_lower(description)) %>% 
  filter(str_detect(description, pattern = 'consumption')) %>% 
  filter(!str_detect(description, pattern = 'total')) %>% 
  mutate(yr_mo_day = lubridate::parse_date_time(yyyymm, "ym")) %>% 
  mutate(month_sep = yearmonth(yr_mo_day)) %>% 
  mutate(value = as.numeric(value)) %>% 
  drop_na(month_sep, value) %>% 
  select(-yyyymm, -msn)

# Create a version with month and year in separate columns:

renew_parsed <- renew_clean %>% 
  mutate(month = lubridate::month(yr_mo_day, label = TRUE)) %>% 
  mutate(year = lubridate::year(yr_mo_day))
```

### **Step 3:** Visualization.

```{r}
renew_gg <-
  ggplot(data = renew_clean,
         aes(x = month_sep,
             y = value)) +
  geom_line(aes(color = description))
  
```

Update colors using paletteer:

```{r}
renew_gg +
  scale_color_paletteer_d("calecopal::bigsur")
```

### **Step 4:** Coerce renew_parsed to a tsibble.

```{r}
renew_ts <- as_tsibble(renew_parsed,
                       key = description,
                       index = month_sep)
```

Look at the data in a few ways:

```{r}
renew_ts %>% autoplot(value)

renew_ts %>% gg_subseries(value)

# renew_ts %>% gg_season(value)

ggplot(data = renew_parsed,
       aes(x = month,
           y = value,
           group = year)) +
  geom_line(aes(color = year)) +
  facet_wrap(~description,
             ncol = 1,
             scales = "free",
             strip.position = "right")
```

Just look at wind energy cons:

```{r}
wind_ts <- renew_ts %>% 
  filter(description == "wind energy consumption")

wind_ts %>% autoplot(value)

wind_ts %>% gg_subseries(value)

ggplot(data = wind_ts,
       aes(x = month,
           y = value,
           group = year)) +
  geom_line(aes(color = year))

```

# Quarterly average consumption for wind:

```{r}
wind_quarter <- wind_ts %>% 
  index_by(year_qu = ~(yearquarter(.))) %>% 
  summarize(avg_cons = mean(value))

head(wind_quarter)
```

### **Step 5:** Decomposition and modeling

```{r}
dcmp <- wind_ts %>% 
  model(STL(value ~season(window = 5)))

components(dcmp) %>% autoplot()
```

```{r}
hist(components(dcmp)$remainder)

# ACF:

wind_ts %>% 
  ACF(value) %>% 
  autoplot()
```

# DANGER DANGER - modeling, know what you're doing (and do a lot of reading)

```{r}
wind_model <- wind_ts %>% 
  model(
    ARIMA(value),
    ETS(value)
  ) %>% 
  fabletools::forecast(h = "10 years")

wind_model %>% autoplot(filter(wind_ts,
                               year(month_sep) > 2010))
```

## Make a world map!

```{r}
world <- read_sf(dsn = here::here('TM_WORLD_BORDERS_SIMPL-0.3-1'),
                 layer = 'TM_WORLD_BORDERS_SIMPL-0.3')
```

